---
title: "Exercise 7 - Nonlinear Models and Tree-based Methods"
author: "Ken Benoit, Slava Mikhaylov, and Jack Blumenau"
output: html_document
---

```{r}
library(ISLR)
library(MASS)
library(class)
library(caret)
set.seed(201785066) #LSE ID
data("College", package = "ISLR")
head(College)
```

### Exercise 7.1

This question relates to the `College` dataset from the `ISLR` package.

(a) Split the data into a training set and a test set. Using out-of-state tuition as the response and the other variables as the predictors, perform appropriate model selection of your choice (from day6) on the training set in order to identify a satisfactory model that uses just a subset of the predictors.

```{r}
trainIndex <- createDataPartition(College$Outstate, p = 0.8, list = FALSE) 
College.train <- College[trainIndex, ]
College.test <- College[-trainIndex, ]
```

```{r}
rr.fitControl <- trainControl(method = "repeatedcv",
                              number = 10,
                              repeats = 10)

rr.model <- train(Outstate ~ .,
                  data = College.train, 
                  method = "foba", 
                  trControl = rr.fitControl,
                  tuneLength = 16)
rr.model
```

```{r}
plot(rr.model)
rr.pred <-  predict(rr.model, newdata = College.test)

mean((College.test[, "Apps"] - rr.pred)^2)
```

```{r}
glmnet.fitControl <- trainControl(method = "cv",
                                  number = 5)

glmnet.fitGrid <- expand.grid(alpha = 1, 
                              lambda = 0:200/10)

glmnet.model <- train(Outstate ~ .,
                      data = College.train, 
                      method = "glmnet", 
                      trControl = glmnet.fitControl,
                      tuneGrid = glmnet.fitGrid,
                      thresh = 1e-12)

glmnet.model
```

```{r}
plot(glmnet.model)
glmnet.pred <-  predict(glmnet.model, newdata = College.test)

mean((College.test[, "Apps"] - glmnet.pred)^2)
```

```{r}
coefficients(glmnet.model$finalModel, s = glmnet.model$bestTune$lambda)
```



(b) Fit a GAM on the training data, using out-of-state tuition as the response and the features selected in the previous step as the predictors. Plot the results, and explain your findings.

```{r}
gam.0 <- gam(Outstate ~ Private + Room.Board, data = College.train)
gam.1 <- gam(Outstate ~ Private + s(Room.Board, df = 2), data = College.train) #Smoothening spline


```

(c) Evaluate the model obtained on the test set, and explain the results obtained.
(d) For which variables, if any, is there evidence of a non-linear relationship with the response?

### Exercise 7.2 

Apply bagging and random forests to a data set of your choice. Be sure to fit the models on a training set and to evaluate their performance on a test set. How accurate are the results compared to simple methods like linear or logistic regression? Which of these approaches yields the best performance?

